---
title: "Hazel Dormouse: Summary of distribution information for Cornwall"
Author: "David J Baker"
output: 
  html_document:
    fig_caption: yes
---

<style>
p.caption {
  font-size: 1.2em;
}
</style>

```{r setup, include=FALSE}

# Set global knitr options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Packages
library(devtools)
library(styler)
library(tidyverse)
library(sf)
library(rgdal)
library(gstat)
library(ggsn)
library(scico)
library(cowplot)
library(gtools)
library(vegan)
library(ranger)

# Load package functions
load_all()

# Shared data path
shareDir <- "C:/Users/DB625/OneDrive - University of Exeter/Fellowship_data/"

# UK shapefile
set_transform_wkt_comment(TRUE)
cornwall_sf <- st_read(paste0(shareDir, "EnglishCountiesCornwall_2011"),"england_ct_2011")

# Set grain
grain <- 4

# hexGrid 'grain' km2
hexGrid_km <- makeHexGrid(cornwall_sf, hexCellArea = grain)

# Extract butterfly data from ERCCIS database
sqlite_file <- paste0(shareDir, "erccis_oct2020.sqlite")
grp <- erccis.db.tbl(sqlite_file, "erccis_oct2020", sppGrps = 'terrestrial mammal')

# Project butterfly data to BNG
grpPrj <- db.transform.proj(grp, cornwall_sf, region = "Cornwall", df = FALSE)

# Filter for target species 
spp <- focal.species.select(grpPrj, "Hazel Dormouse", cornwall_sf, grain, hexGrid = hexGrid_km, collapse_time = FALSE)


```

## Overview of Data

### Distribution of records 2000-2019

The distribution of dormouse records in the ERCCIS database for the years 2000 - 2019 (inclusive). 

```{r, echo = FALSE, fig.cap = "Figure 1. Distribution of Hazel Dormouse records (2000 - 2019)"}

species.evaluation(grpPrj, "Hazel Dormouse", 2000, 2019, cornwall_sf, plotType = 'Map') 


```

### Number of records per year 2000-2019

```{r, echo = FALSE, fig.cap = "Figure 2. Number of Hazel Dormouse records per year"}

species.evaluation(grpPrj, "Hazel Dormouse", 2000, 2019, cornwall_sf, plotType = 'Hist') 


```

## Summary of sampling bias and effort

We do not have non-detection data in the database and therefore have no data on sites that have been surveyed for dormice but non were detected. Thus, we do not know the degree to which these data are spatially biased and must infer this information from other data. One way of estimating bias is to look at the spatial distribution of records across similar species and use this information to infer the distribution of sampling effort across the landscape. 

One way to map the degree of effort is to calculate the degree of inventory completeness in a cell and then map this variation across the region.

```{r, echo = FALSE}

invComp <- inventory.completeness(grpPrj, "terrestrial mammal", start_yr = 2000, end_yr = 2019, 1, cornwall_sf, grain, hexGrid = hexGrid_km)

sampComp <- invComp %>% 
  dplyr::filter(year == 2019) %>%
  mutate(coverage = obs_spp / chao, site = as.character(site)) %>%
  left_join(hexGrid_km, . , by = c("id" = "site")) %>%
  ggplot() + 
  geom_sf(aes(fill = coverage)) +
  scale_fill_scico() + theme_minimal()
print(sampComp)

#tg <- target.group.weights(grpPrj, "terrestrial mammal", cornwall_sf, grain, hexGrid = hexGrid_km, showPlot = TRUE) 
scico_palette_names()
```

# Simulation example

### Environmental data

CEH Landcover 2019 1km
*LC_1 = Deciduous woodland
*LC_3 = Arable
*LC_4 = Improved grassland
*LC_7 = Acid grassland
*LC_9 = Heather
*LC_10 = Heather grassland
*LC_11 = Bog
*LC_14 = Freshwater
*LC_15 = Supralittoral rock
*LC_16 = Supralittoral sediment
*LC_17 = Littoral rock
*LC_18 = Littoral sediment
*LC_19 = Saltmarsh
*LC_20 = Urban
*LC_21 = Suburban

Microclimate data
*
* etc

```{r, echo = FALSE}

# Load CEH land use data at 1 km2 with buffer 0 and 1
land_cover_c19 <- get(load("C:/Users/DB625/OneDrive - University of Exeter/Fellowship_data/cornwall_ceh2019_landcover_1km.Rdata"))
land_cover_c19 <- land_cover_c19[,c('id', 'X', 'Y', 'grain', paste0('LC_', c(1,3,4,7,9,10,11,14,15,16,17,18,19,20,21),'_0'))]

# Load bioclimate variables
bioClim <- get(load("C:/Users/DB625/OneDrive - University of Exeter/Fellowship_data/cornwall_climvars_1km_means.Rdata"))
names(bioClim)

# Merge 
enviro_dat <- left_join(land_cover_c19, bioClim, by = c('id', 'X', 'Y', 'grain'))
enviro_vars <- names(enviro_dat)[grep(names(enviro_dat), pattern = 'LC|_m')]

```

### Simulate a 50 species assemblage and sample

```{r, echo = FALSE}

# Simulate assemblage
spp50 <- occupancy.multi(enviro_dat, enviro_vars, n_spp = 50, pca_axes = 3, 
                         cornwall_sf, plotRichness = TRUE)

# Extract survey data
surv50 <- virtual.sampling(spp50, p_sites = 0.20, n_yrs = 10, spat_bias = TRUE, 
                           samp_weights = NULL, bias_level = 3, rep_visit = 2.5, 
                           maxVisit = 10, reli_score =  "high")

# Example species plots
plot.virtual.species(spp50, surv50, 20, 1:10)

# Extract data for one species and models encounter rate

# Zero filling equivalent lists
focSpp <- surv50 %>%
  dplyr::select(id, X, Y, det_2) %>%
  dplyr::group_by(id, X, Y) %>%
  dplyr::summarise(det = max(det_2)) %>%
  left_join(enviro_dat, by = c("id", "X", "Y")) %>%
  dplyr::select(id, X, Y, det, LC_1_0, LC_19_0, LC_20_0, LC_21_0, temp_ann_m, temp_gs_m, 
               mtcm_m, mtcq_m, frost_m, precip_gs_m, soilm_gs_m) %>%
  drop_na()

focSpp$det
# Split to training and test data
focSpp$set <- rbinom(nrow(focSpp), 1, 0.25)
sim.test <- focSpp[focSpp$set == 1,]
sim.fit <- focSpp[focSpp$set == 0,]

# Random forest
form <- formula(det ~ LC_1_0 + LC_19_0 + LC_20_0 + LC_21_0 + temp_ann_m + temp_gs_m + 
               mtcm_m + mtcq_m + frost_m + precip_gs_m + soilm_gs_m)
rf <- ranger(form, data = sim.fit, importance = "permutation")

pred <- predict(rf, data = sim.test)
rbind(sim.test$det, predictions(pred))
importance(rf)

```

```{r, echo = FALSE}

#~#~#
#~#~# SDM
#~#~#
#~# Focal species
foc_sp_samp <- na.omit(vs_survey[[sp]])
foc_sp_samp$w <- 1
n <- length(foc_sp_samp$pa[foc_sp_samp$pa==1])
foc_cell <- cellFromXY(bk,foc_sp_samp[,1:2])

#~# Prevalence
foc_sp <- na.omit(vs_com[[sp]])
prev <- length(foc_sp$pa[foc_sp$pa==1])/length(foc_sp$pa)

#~# Background sample
bkSmap <- bkMap[!(bkMap$id %in% foc_cell),]
bkSamp <- sample_n(bkSmap,2000,weight=bkSmap$bk)
bkSamp$w <- ((n*(1-prev))/prev)/2000
bkSamp$pa <- 0
plot(rasterFromXYZ(bkSamp))

#~# 
spDat <- rbind(foc_sp_samp[,c('x','y','pa','w')],bkSamp[,c('x','y','pa','w')])
spDat <- na.omit(cbind(spDat,raster::extract(swClimRast,spDat[,1:2])))

# PCA covariates
pca <- prcomp(na.omit(raster::extract(swClimRast,vs_com[[sp]][,1:2])))
clim <- na.omit(cbind(vs_com[[20]],raster::extract(swClimRast,vs_com[[sp]][,1:2])))
pca_clim <- cbind(clim[,1:2],predict(pca, newdata=clim))
pca_clim_rst <- rasterFromXYZ(pca_clim)
spDat <- na.omit(cbind(spDat,raster::extract(pca_clim_rst,spDat[,1:2])))

# Models
m1 <- glm(pa ~ .^3,family=binomial(link = "logit"),data=spDat[,c('pa','PC1','PC2','PC3','PC4','PC5')],weights=spDat$w)

# predict to the new dataset
pca_clim$pred <- predict(object=m1, newdata=pca_clim, type='response')
pca_clim_r <- rasterFromXYZ(pca_clim[,c('x','y','pred')])
w <- gaussWindow(4 * 5, 4)
pca_clim_dens.ras <- focal(pca_clim_r, w = w / sum(w),pad = TRUE,na.rm = TRUE)
pca_clim_dens.ras <- mask(pca_clim_dens.ras,pca_clim_r)
pca_clim_dens.ras <- na.omit(data.frame(coordinates(pca_clim_dens.ras),pred=getValues(pca_clim_dens.ras)))

sdm_plot <- ggplot() +geom_raster(aes(x=x,y=y,fill=pred),data=pca_clim_dens.ras) + 
  scale_fill_distiller(palette = 'YlGnBu',direction = 1,name='P(1)') + 
  #geom_point(aes(x=x,y=y),size=0.5,data=vs_com[[2]][vs_com[[2]]$pa==1,])+
  coord_cartesian() + theme_minimal() %+replace% 
  theme(axis.title = element_blank(),axis.text = element_blank(),legend.position = c(0.1,0.7))
sdm_plot
ggsave('Outputs/SDM_projection_figure.png',plot=sdm_plot,width=6,height=4)



```
